
##Lab1 Answers File
--
Please write your answers into the space below the questions. Knit this file when you are done to check that all code runs. Remember to rename the file to contain *your actual* Last and First name.
--

<span style="color:green"> 3.1</span>
```{r}
# Calculate pi to the power of 6
pi^6
```


<span style="color:green"> 3.2.</span>
```{r}
# Calculate 3 million times pi
3e6*pi
```


<span style="color:green"> 4.1</span>

<span style="color:blue"> 
a) R treats TRUE as 1 and FALSE as 0
</span>

<span style="color:blue"> 
b) We may sometimes want to run a trial with 2 outcomes (TRUE/FALSE) and maybe count the number of successful trials. In R, to do this we can just sum all outcomes. 
</span>


<span style="color:green"> 4.2</span>

<span style="color:blue"> 
a) y is actually storing a value. y is unchanged when we alter x.
</span>

<span style="color:blue">
b) If a variable stores an expression, it will be changed if one of the variables within the expression changes. On the other hand, if a variable just holds a value, it will not change
</span>


<span style="color:green"> 4.3</span>

<span style="color:blue"> 
The function c() combines its parameters into a vector and returns it.
</span>


<span style="color:green"> 4.4</span>
```{r}
# Subtract 8 from myVec
myVec <- c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
newVec <- myVec - 8
newVec
```
<span style="color:blue"> 
This operation subtracts 8 from each element of myVec. 
</span>

<span style="color:green"> 4.5</span>
```{r}
# Calculate the mean of myVec
mean(myVec)
```

<span style="color:green"> 4.6</span>
```{r}
#myVec[3:5]
#myVec[-3]
#myVec[c(1,3,8)]  
#2*myVec 
#myVec[1] <- 7.5
```

<span style="color:green"> a.) `myVec[3:5]` <br> </span>

<span style="color:blue"> 
Selects the third through fifth (inclusive) elements from myVec 
</span>

<span style="color:green"> b.) `myVec[-3]` <br> </span>

<span style="color:blue"> 
Removes the third element from myVec 
</span>

<span style="color:green"> c.) `myVec[c(1,3,8)]` <br> </span>
<span style="color:blue"> 
Selects the first, third, and eighth element from myVec 
</span>

<span style="color:green"> d.) `2*myVec`  <br> </span>

<span style="color:blue"> 
Multiplies each element of myVec by 2 
</span>

<span style="color:green"> e.) `myVec[1] <- 7.5` <br> </span>

<span style="color:blue"> 
Sets the first element of myVec to 7.5
</span>


<span style="color:green"> 4.7 <br> </span>
<span style="color:blue"> a) </span>
```{r}
# Create vector of heart rates from 60 to 100 in increments of 2
heartRate <- seq(60, 100, by = 2)
heartRate
```
<span style="color:blue"> b) </span>
```{r}
# create vector ejectionVolumes containing the total amount of blood (in mL) ejected in 30 minutes for each heart rate in your range (Assuming ejection rate of 69 mL/beat)

averageEjection <- 69
# number of beats in 30 min
beatsTotal <- heartRate * 30

# total volume ejected 
ejectionVolumes <- beatsTotal * averageEjection
ejectionVolumes
```


<span style="color:green"> 5.1 <br> </span>
<span style="color:blue">
"!=" means not equals to. ">=" is greater than or equal to. "<=" is less than or equal to. "==" is equals to. "<" is strictly less than. ">" is strictly greater than
</span>


<span style="color:green"> 5.2 <br> </span>
<span style="color:blue">
"||" is the OR operator
</span>

<span style="color:green"> 5.3 <br> </span>
<span style="color:blue">
The operation is boolean indexing into the vector multiplesOfFive and selecting the entries that are not divisible by ten.
</span>

<span style="color:green"> 6.1</span>
```{r}
carSpeed <- 65
speedLimit <- 60
if(carSpeed > speedLimit) { 
    print("Give ticket") 
} else { 
    print("Do nothing") 
} 
```
<span style="color:blue">
a) If the value stored in carSpeed is greater than that stored in speedLimit, then the code prints "Give ticket", otherwise "Do nothing".
</span>

<span style="color:blue">
b) The carSpeed is still greater than the speedLimit, so the code still prints out "Give ticket".
</span>

<span style="color:green"> 6.2</span>
```{r}
if(1) {
    print("a")
} else {
    print("b")
}
```
<span style="color:blue">
1 is interpreted as TRUE, so the code within the if statement runs.
</span>


<span style="color:green"> 6.3</span>
```{r}
# Find the minimum of 3 numbers
x <- 4
y <- 5
z <- 3
if(x < y){
  if(x < z){
    print(x)
  }
  else{
    print(z)
  }
} else {
  if (y < z){
    print(y)
  }
  else {
    print(z)
  }
}
```

<span style="color:green"> 6.4</span>
```{r}
# instructs patients to “seek medical attention” if their bodyTemp is above 45 degrees or below 40 degrees, or to “do nothing” if their temperature is within an acceptable range.

bodyTemp = 43

if (bodyTemp > 45 || bodyTemp < 40){
  print("seek medical attention")
} else {
  print("do nothing")
}
```



<span style="color:green"> 7.1</span>

<span style="color:blue"> 
a) "i" us defined within the for loop. It is implicitly defined at the beginning of the for loop, and then reassigned each iteration of the loop <br>
</span>

<span style="color:blue"> 
b) "i" is a global variable because it shows up in the Global Environment <br>
</span>

<span style="color:blue"> 
c) The loop is printing out the first 10 even numbers <br>
</span>

<span style="color:blue"> 
d) Defining "i" outside of the loop before its execution does not affect the calculation. <br>
</span>


<span style="color:green"> 7.2 <br> </span>
<span style="color:blue">
a) If x %% 2 == 0, then x is even. If x %% 2 == 1, then x is odd. <br>
</span>

<span style="color:blue">
b)
</span>
```{r}
#for loop that iterates from 1 to 10, and prints out the even numbers between 1 and 10 as well as the number 5.
for (i in 1:10){
  if (i %% 2 == 0 || i == 5){
    print(i)
  }
}
```


<span style="color:green"> 7.3</span>
```{r}
#for loop that iterates over letters and prints each letter in upper-case
for (i in 1:length(letters)){
  print(toupper(letters[i]))
}
```


<span style="color:green"> 7.4 <br> </span>
<span style="color:blue">
a) It would result in an infinite loop because the variable x is never updated so it (x = 3) will always be less than or equal to 6 <br>
</span>
<span style="color:blue">
b)
</span>
```{r}
#x <- 3  
#while(x <=  6){  
#   print(x + x)  
#} 

# modified so loop will be functional
x <- 3
while (x <= 6){
  print(x + x)
  x <- x + 1
}
```

<span style="color:green"> 8.1 </span>

<span style="color:blue">
a) <br>
</span>
```{r}
# function that takes in one parameter n and returns its square
findSquare <- function(n){
  return(n^2)
}
```
<span style="color:blue">
b) <br>
</span>
```{r}
# calculate square of 269 and silently assign it to variable squared
squared <- findSquare(269)
```


<span style="color:green"> 8.2 <br> </span>
<span style="color:blue">
The provided inputs are assigned to the function parameters in the same order as provided. If the function takes in x, y, z in that order, then the first input will be assigned to x, second to y, and third to z. Alternatively, you can call the function and explicitly define the parameters (ex. passing in x = 3) and then the order in which you input doesn't matter.
</span>


<span style="color:green"> 8.3 </span>

<span style="color:blue"> a) </span>
```{r}
# calculates mean of vector
calcMean <- function(vec){
  return(sum(vec)/length(vec))
}
```
<span style="color:blue"> b) </span>
```{r}
# execute calcMean on vector containing values from 9 to 30
calcMean(9:30)
```
<span style="color:blue"> c) </span>
```{r}
# check using Rs built in mean function
mean(9:30)
```



<span style="color:green"> 8.4 <br> </span>

<span style="color:blue"> a) </span>
```{r}
# Using a for loop convert each pH from 1-10 into concentrations of H+. Store the list of [H+] in a vector.
myVec <- numeric(10)
for (ph in 1:10){
  myVec[ph] <- 10^(-ph)
}
myVec
```
<span style="color:blue"> b) </span>
```{r}
# function called H_from_pH that, given a pH value, outputs the concentration of H+.
# function called OH_from_pH that, given a pH value, outputs the concentration of OH-
H_from_pH <- function(pH){
  return(10^(-pH))
}

OH_from_pH <- function(pH){
  return(10^(-(14-pH)))
}
```

<span style="color:green"> 8.5 <br> </span>
<span style="color:blue"> a) </span>
```{r}
# function that returns the standard deviation of a vector of numbers
calc_std <- function(vec){
  mu <- mean(vec)
  return((sum((vec-mu)^2)/(length(vec)-1))^0.5)
}
```
<span style="color:blue"> b) </span>
```{r}
# function executed on vector 1-200 with step size of 0.25
calc_std(seq(1, 200, by=0.25))
```
<span style="color:blue"> c) </span>
```{r}
# confirm results of function match R's built in standard deviation function
sd(seq(1, 200, by=0.25))
```


